Long range correlations and their event scale dependence in high multiplicity pp collisions at $\sqrt{s} = 13$~TeV

### 2020-06-23 sec_ana.tex
update on sec 5.2
adopted jasper's comment to Sec3

The V0 detector consists of two rings, V0-A and V0-C, each made of 32 scintillator tiles, covering the full azimuthal angle within $2.8 < \eta < 5.1$ and $-3.7 < \eta < -1.7$, respectively. The V0 provides trigger and estimation of event multiplicity. A sample of events including higher numbers of produced particles is obtained with a high multiplicity trigger, which is achieved by requiring some thresholds on signal amplitudes of energy deposit by charged particles in the V0 detector.

Charged particles are reconstructed by the ITS and TPC, which are working in a constant solenoidal magnetic field of 0.5~T.  The ITS is composed of three sub-systems, Silicon Pixel Detector (SPD), Silicon Drift Detector (SDD) and Silicon Strip Detector (SSD). The ITS and TPC covering the full azimuthal region have acceptances up to $|\eta| < 1.4$ and 0.9, respectively, for detection of charged particles with a primary vertex in $|z_\mathrm{vtx}| < 10$~cm. The tracking of charged particles is done with the combined information of the ITS and TPC that enables the reconstruction of tracks down to 0.2~GeV/$c$ with a $\sim $75\% efficiency.

@@

The yield of two-particle correlations between the trigger and associated particles are measured as functions of relative pseudorapidity ($\Delta\eta$) and azimuthal angle differences($\Delta\varphi$)
\begin{eqnarray}
\frac{1}{N_{\rm{trig}}} \frac{ \rm{d}\it{}^{2} N_{\rm{pair}} }{ \rm{d} \Delta\eta \rm{d}\Delta\varphi} = B(0, 0)\frac{S(\Delta\eta, \Delta\varphi)}{B(\Delta\eta, \Delta\varphi)}\quad,
\label{eq:corrfunction}
\end{eqnarray}
where  $\pttrig$ and $\ptassoc$ are the transverse momenta of  the trigger and associated particles, respectively, $N_\mathrm{trig}$ is the number of trigger particles, $N_\mathrm{pair}$ is the number of pairs of the trigger and associated particles, and $S (\Delta\eta, \Delta\varphi)$ and $B (\Delta\eta, \Delta\varphi)$ are the signal and background distributions, respectively. $S (\Delta\eta, \Delta\varphi)$ is constructed using two-particle correlations in the same event and $B(\Delta\eta, \Delta\varphi)$ is the yield of two-particle correlations in mixed events. The background distribution is normalized by its value at $\Delta\eta$ and $\Delta\varphi = 0$, represented as $B (0,0)$. In mixed events, the primary vertices and multiplicity classes are required to be in the same range to avoid the acceptance effect with respect to the multiplicity class and primary vertex.

The ridge yields at large $\Delta\eta$ are extracted in different multiplicity classes and transverse momentum intervals. The large $\Delta\eta$ range is selected as $1.6<|\Delta\eta|<1.8$ where non-flow effects (mainly coming from jets) are negligible. In this region, the per trigger yield is expressed as
\begin{eqnarray}
\frac{1}{N_{\rm{trig}}} \frac{ \rm{d}\it{}N_{\rm{pair}} }{ \rm{d}\Delta\varphi } = \int_{|\Delta \eta|>1.6} \rm{d} \Delta \eta \frac{1}{\it{N}_{\rm{trig}}} \frac{ \rm{d}\it{}^{2} N_{\rm{pair}} }{ \rm{d}\Delta\eta d\Delta\varphi} \quad.
\end{eqnarray}

The baseline of the correlations is subtracted by implementing Zero-Yield-At-Minimum (ZYAM) procedure~\cite{Ajitanand:2005jj}. The minimum yield $(C_{\rm{ZYAM}})$ at the $\Delta\varphi$ ($\Delta\varphi_{\rm{min}}$) in the $\Delta\varphi$ distribution is obtained from the function, which fits the $\Delta\varphi$ distribution with Fourier series up to the third harmonic. Subtracting $C_{\rm{ZYAM}}$ from the $\Delta\varphi$ distribution makes the yield at $\Delta\varphi_{\rm{min}}$ zero in order to describe the shape to find the minimum to be used for integral of the yield. The associated yield $Y^{\rm{assoc}}$ (ridge yield is obtained by integrating the near-side peak of the $\Delta\varphi$ distribution over $|\Delta\varphi|<|\Delta\varphi_{\rm{min}}|$ after the ZYAM procedure
\begin{eqnarray}
Y^{\rm{assoc}} = \int_{|\Delta \varphi| < |\Delta\varphi_{\rm{min}}| } \rm{d} \Delta\varphi \frac{1}{\it{N}_{\rm{trig}}} \frac{ \rm{d}\it{}N_{\rm{pair}} }{ \rm{d}\Delta\varphi } \quad.
\end{eqnarray}

The ridge yield is further studied by exploiting the momentum transfer between the interacting partons, so called, event scale. On average, increasingly hard parton interactions result from pp collisions with decreasing impact parameters between the two protons~\cite{Sjostrand:1986ep,Frankfurt:2010ea}.
In this article, the event scales are set by requiring minimum transverse momentum of leading track or jet reconstruction in mid-rapidity since leading tracks with a high $p_\mathrm{T}$ or jets result from the hard scattering. The leading track is selected within $|\eta|<0.9$ and the full azimuthal angle. Jets are reconstructed with charged particles only (charged jets) for the anti-$k_{\rm{T}}$ algorithm and the resolution parameter $R = 0.4$. Jets are selected in $|\eta_\mathrm{jet}|<0.4$ and the full azimuthal angle. The transverse momentum of jets $p_\mathrm{T,\,jet}$ is corrected for underlying event density that is measured for the $k_{\rm{T}}$ algorithm with $R = 0.2$.

Monte Carlo simulation with $\pythiam$ event generator and particle transport inside ALICE using GEANT 3 is used to correct the acceptance and the efficiency of the ALICE detectors.

@@

\section{Systematic Uncertainty}
\label{sec:uncertainties}
The systematic uncertainty is evaluated by varying the event selection, track selection, yield extraction procedures and detector efficiency correction. Each source is assumed uncorrelated. The uncertainty from the event selection can be divided into pile-up rejection and primary vertex selection. The uncertainty from the track selection is estimated by changing the selection criteria. The uncertainty from the yield extraction procedures can be categorized into long-range definition and ZYAM procedure. The uncertainty from the efficiency correction is estimated by enlarging the event mixing bins, which is relevant to the acceptance correction.
%Finally, The uncertainty from the application of the detector efficiency correction is estimated comparing the results constructed by efficiency corrected tracks with the results constructed. 
Finally, the uncertainty from the efficiency correction for charged tracks is estimated by comparing the correlations function constructed by reconstructed tracks from the detector responses with the correlation function constructed by generated particles from the event generator.
The uncertainty of $Y^{\rm{assoc}}(\pttrig,\,\ptassoc)$ is summarized by averaging each uncertainty for given $\pttrig$ and $\ptassoc$. The uncertainty of $Y^{\rm{assoc}}(\ptlead)$ and $Y^{\rm{assoc}}(\ptjet)$ is summarized by averaging the each uncertainty for a given $\ptlead$ and $\ptjet$ selection, respectively.
\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|}
\hline 
\multirow{2}{*}{Sources}  & \multicolumn{3}{c|}{Systematic uncertainty (\%)} \\\cline{2-4}
         & $Y^{\rm{assoc}}(\it{p}_{\rm{T, trig}},\,\it{p}_{\rm{T, assoc}})$ & $Y^{\rm{assoc}}(\it{p}_{\rm{T, Lead}})$ & $Y^{\rm{assoc}}(\it{p}_{\rm{T, Jet}})$ \\ \hline \hline
Pileup cut			& $\sim$0.8	&0--4		&0--4		\\ \hline
Primary Vertex		& $\sim$2.4	&1--12	&1-8		\\ \hline

@@

The uncertainty from the pileup rejection is estimated by measuring the changes of results with different rejection criteria from the default one. The pileup rejection is achieved by reconstructing the multiple vertex in the individual event. The multiplicity of the reconstructed vertices determines whether the event is contaminated by pileup or not. The variation is set by the minimum multiplicity of reconstructed vertices from 5 to 3. 
%Pileup cut (multbins) -> (default pileup from the physics selection)
%Pileup cut (multbins) -> PileupMV (?)
The estimated uncertainty is 1\% without $\ptlead$ or $\ptjet$ selection. The uncertainty is increased up to 4\% with $\ptlead$ or $\ptjet$ selection. The uncertainty from the primary vertex selection along the beam axis is estimated by varying the selection range of primary vertex from $|z_\mathrm{vtx}|<$ 8 to 6 cm. The narrower primary vertex selection allows one to test the acceptance effect on the measurement. The estimated uncertainty is increased up to 12\% from 1\%.
The uncertainty from the track selection is estimated by varying the track selection criteria from the one optimized for refining the angular track distribution to the other optimized for particle identification. The variation is achieved by changing the mandatory SPD requirement and the maximal threshold for chi-square of the track fit.
%https://github.com/alisw/AliRoot/blob/master/ANALYSIS/ESDfilter/macros/AddTaskESDFilter.C
The estimated uncertainty for track selection is 4\% by averaging the uncertainty for whole transverse momentum bins and 2\% for $1<\it{p}_{\rm{T}}<2$ GeV/$c$ tracks.

The uncertainty from the ZYAM procedure is estimated by varying the finding-scope for minimum position. The estimated uncertainty for track selection is 5.1\% by averaging the uncertainty for the whole transverse momentum bins and 2\% for $1<\it{p}_{\rm{T}}<2$ GeV/$c$ tracks, which is not largely affected by the event scale selection. The uncertainty from the $\Delta\eta$ projection range is estimated by varying the $\Delta\eta$ range. The uncertainty might include effects from the longitudinal de-correlations and possible non-flow contamination, which is thought be small. The estimated uncertainty is increased up to 9\% from 3\%, which is obtained by averaging the whole transverse momentum range.
The uncertainty from the mixing pool bin size of primary vertex is estimated by varying the bin size from 2 to 1 cm.

>>>>

The V0 detector consists of two rings, V0-A and V0-C, each made of 32 scintillator tiles, covering the full azimuthal angle within $2.8 < \eta < 5.1$ and $-3.7 < \eta < -1.7$, respectively. The V0 provides trigger and estimation of event multiplicity. A sample of events including higher numbers of produced particles is obtained with a high multiplicity trigger, which is achieved by requiring 5 times of mean signal amplitudes of energy deposit by charged particles in the V0 detector.

Charged particles are reconstructed by the ITS and TPC, which are working in a constant solenoidal magnetic field of 0.5~T. The ITS is composed of three sub-systems, Silicon Pixel Detector (SPD), Silicon Drift Detector (SDD) and Silicon Strip Detector (SSD). The ITS and TPC covering the full azimuthal region have acceptances up to $|\eta| < 1.4$ and 0.9, respectively, for detection of charged particles with a primary vertex in $|z_\mathrm{vtx}| < 8$~cm. The tracking of charged particles is done with the combined information of the ITS and TPC that enables the reconstruction of tracks down to 0.15~GeV/$c$ with a $\sim $65\% efficiency.
%https://arxiv.org/pdf/1910.14400.pdf

@@

The two-particle correlations function between the trigger and associated particles are measured as functions of relative pseudorapidity ($\Delta\eta$) and azimuthal angle differences($\Delta\varphi$)
\begin{eqnarray}
\frac{1}{N_{\rm{trig}}} \frac{ \rm{d}\it{}^{2} N_{\rm{pair}} }{ \rm{d} \Delta\eta \rm{d}\Delta\varphi} = B(0, 0)\frac{S(\Delta\eta, \Delta\varphi)}{B(\Delta\eta, \Delta\varphi)}\quad,
\label{eq:corrfunction}
\end{eqnarray}
where  $\pttrig$ and $\ptassoc$ are the transverse momenta of  the trigger and associated particles, respectively, $N_\mathrm{trig}$ is the number of trigger particles, $N_\mathrm{pair}$ is the number of pairs of the trigger and associated particles, and $S (\Delta\eta, \Delta\varphi)$ and $B (\Delta\eta, \Delta\varphi)$ are the signal and background distributions, respectively. $S (\Delta\eta, \Delta\varphi)$ is constructed using two-particle correlations in the same event and $B(\Delta\eta, \Delta\varphi)$ is in mixed events. The background distribution is normalized by its value at $\Delta\eta$ and $\Delta\varphi = 0$, represented as $B (0,0)$. In mixed events, the primary vertices and multiplicity classes are required to be in the same range to avoid the acceptance effect with respect to the multiplicity class and primary vertex.

The ridge yields at large $\Delta\eta$ are extracted in different multiplicity classes and transverse momentum intervals. The large $\Delta\eta$ range is selected as $1.6<|\Delta\eta|<1.8$ where non-flow effects (mainly coming from jets) are negligible. In this region, the $\Delta\varphi$ distribution, or the so called per-trigger yield is expressed as
\begin{eqnarray}
\frac{1}{N_{\rm{trig}}} \frac{ \rm{d}\it{}N_{\rm{pair}} }{ \rm{d}\Delta\varphi } = \int_{|\Delta \eta|>1.6} \rm{d} \Delta \eta \frac{1}{\it{N}_{\rm{trig}}} \frac{ \rm{d}\it{}^{2} N_{\rm{pair}} }{ \rm{d}\Delta\eta d\Delta\varphi} \quad.
\end{eqnarray}

The baseline of the correlations is subtracted by implementing Zero-Yield-At-Minimum (ZYAM) procedure~\cite{Ajitanand:2005jj}. The minimum yield $(C_{\rm{ZYAM}})$ at the $\Delta\varphi$ ($\Delta\varphi_{\rm{min}}$) in the $\Delta\varphi$ distribution is obtained from the function, which fits the $\Delta\varphi$ distribution with Fourier series up to the third harmonic. Subtracting $C_{\rm{ZYAM}}$ from the $\Delta\varphi$ distribution makes the yield at $\Delta\varphi_{\rm{min}}$ zero in order to describe the shape to find the minimum to be used for integral of the yield. The ridge yield ($Y^{\rm{assoc}}$) is obtained by integrating the near-side peak of the $\Delta\varphi$ distribution over $|\Delta\varphi|<|\Delta\varphi_{\rm{min}}|$ after the ZYAM procedure
\begin{eqnarray}
Y^{\rm{assoc}} = \int_{|\Delta \varphi| < |\Delta\varphi_{\rm{min}}| } \rm{d} \Delta\varphi \frac{1}{\it{N}_{\rm{trig}}} \frac{ \rm{d}\it{}N_{\rm{pair}} }{ \rm{d}\Delta\varphi } \quad.
\end{eqnarray}

The ridge yield is further studied by exploiting the momentum transfer between the interacting partons, so called, event scale. On average, increasingly hard parton interactions result from pp collisions with decreasing impact parameters between the two protons~\cite{Sjostrand:1986ep,Frankfurt:2010ea}.
In this article, the event scales are set by requiring minimum transverse momentum of leading track or reconstructed jet in mid-rapidity since leading tracks with a high $p_\mathrm{T}$ or jets result from the hard scattering. The leading track is selected within $|\eta|<0.9$ and the full azimuthal angle. Jets are reconstructed with charged particles only (charged jets) for the anti-$k_{\rm{T}}$ algorithm and the resolution parameter $R = 0.4$. Jets are selected in $|\eta_\mathrm{jet}|<0.4$ and the full azimuthal angle. The transverse momentum of jets $p_\mathrm{T,\,jet}$ is corrected for underlying event density that is measured for the $k_{\rm{T}}$ algorithm with $R = 0.2$.

Monte Carlo simulation with $\pythiam$ event generator and particle transport inside ALICE using GEANT3~\cite{Brun:1994aa} is used to correct the acceptance and the efficiency of the ALICE detectors.

@@

\section{Systematic Uncertainty}
\label{sec:uncertainties}
The systematic uncertainty is evaluated by varying the event selection, track selection, yield extraction procedures and detector efficiency correction. Each source is assumed to be uncorrelated. The uncertainty from the event selection can be divided into pile-up rejection and primary vertex selection. The uncertainty from the track selection is estimated by changing the selection criteria. The uncertainty from the yield extraction procedures can be categorized into long-range definition and ZYAM procedure. The uncertainty from the efficiency correction is estimated by enlarging the event mixing bins, which is relevant to the acceptance correction.
%Finally, The uncertainty from the application of the detector efficiency correction is estimated comparing the results constructed by efficiency corrected tracks with the results constructed. 
Finally, the uncertainty from the efficiency correction for charged tracks is estimated by comparing the correlations function constructed by reconstructed tracks from the detector responses with the correlation function constructed by generated particles from the event generator.
The uncertainty of $Y^{\rm{assoc}}(\pttrig,\,\ptassoc)$ is summarized by averaging each uncertainty for given $\pttrig$ and $\ptassoc$. The uncertainty of $Y^{\rm{assoc}}(\ptlead)$ and $Y^{\rm{assoc}}(\ptjet)$ is summarized by averaging the each uncertainty for a given $\ptlead$ and $\ptjet$ selection, respectively.
\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|}
\hline 
\multirow{2}{*}{Sources}  & \multicolumn{3}{c|}{Systematic uncertainty (\%)} \\\cline{2-4}
         & $Y^{\rm{assoc}}(\it{p}_{\rm{T, trig}},\,\it{p}_{\rm{T, assoc}})$ & $Y^{\rm{assoc}}(\it{p}_{\rm{T, Lead}})$ & $Y^{\rm{assoc}}(\it{p}_{\rm{T, Jet}})$ \\ \hline \hline
Pileup cut			& $\sim$0.8	&0--4		&0--4		\\ \hline
Primary Vertex		& $\sim$2.4	&1--12	&1-8		\\ \hline

@@

The uncertainty from the pileup rejection is estimated by measuring the changes of results with different rejection criteria from the default one. The pileup rejection is achieved by reconstructing the multiple vertex in the individual event. The multiplicity of the reconstructed vertices determines whether the event is contaminated by pileup or not. The variation is set by the minimum multiplicity of reconstructed vertices from 5 to 3. 
%Pileup cut (multbins) -> (default pileup from the physics selection)
%Pileup cut (multbins) -> PileupMV (?)
The estimated uncertainty is 1\% without $\ptlead$ or $\ptjet$ selection. The uncertainty is increased up to 4\% with $\ptlead$ or $\ptjet$ selection. The uncertainty from the primary vertex selection along the beam axis is estimated by varying the selection range of primary vertex from $|z_\mathrm{vtx}|<$ 8 to 6 cm. The narrower primary vertex selection allows one to test the acceptance effect on the measurement. The estimated uncertainty is increased up to 12\% from 1\%.
The uncertainty from the track selection is estimated by varying the track selection criteria from the one optimized for refining the angular track distribution to the other optimized for particle identification. The variation is achieved by changing the mandatory SPD requirement and the maximal threshold for chi-square of the track fit.
%https://github.com/alisw/AliRoot/blob/master/ANALYSIS/ESDfilter/macros/AddTaskESDFilter.C
The estimated uncertainty for track selection is 4\% by averaging the uncertainty for whole transverse momentum bins and 2\% for $1<\it{p}_{\rm{T}}<2$ GeV/$c$ particles.

The uncertainty from the ZYAM procedure is estimated by varying the finding-scope for minimum position. The estimated uncertainty for track selection is 5.1\% by averaging the uncertainty for the whole transverse momentum bins and 2\% for $1<\it{p}_{\rm{T}}<2$ GeV/$c$ tracks, which is not largely affected by the event scale selection. The uncertainty from the $\Delta\eta$ projection range is estimated by varying the $\Delta\eta$ range. The uncertainty might include effects from the longitudinal de-correlations and possible non-flow contamination, which is thought be small. The estimated uncertainty is increased up to 9\% from 3\%, which is obtained by averaging the whole transverse momentum range. The uncertainty from the mixing pool bin size of primary vertex is estimated by varying the bin size from 2 to 1 cm. The estimated uncertainty is increased up to 8-16\% from 1-2\%.

### 2020-06-23 sec_intro.tex
add refereces and update on ana
adopted jasper's comment to Sec3

Measurements of two-particle angular correlations are typically performed in terms of two dimensional $\Delta\eta$--$\Delta\varphi$ correlation functions, where $\eta$ is the pseudorapidity and $\varphi$ is the azimuthal angle. Long-range structure of two-particle correlation functions is of particular interest in studies of possible novel collective effects, where the effects of known sources such as resonance decays and fragmentation of high-momentum partons are known to be small. For most Monte Carlo (MC) event generators in pp collisions, the typical sources of such long-range correlations are momentum conservation and away-side ($\Delta\varphi$ $\approx$ $\pi$ ) jet correlations.
The enhancement in the associated yield of two-particle correlations at small relative azimuthal angle ($\Delta\varphi$) that extends over a long-range of relative pseudorapidity ($\Delta\eta$), is dubbed ``ridge'' due to its shape in the $\Delta\eta$--$\Delta\varphi$ plane.
The shape of these $\Delta\varphi$ correlations can be studied via a Fourier decomposition~\cite{Poskanzer:1998yz,Voloshin:2008dg}. The second and third order terms are the dominant harmonic coefficients $v_n$. The $v_n$ coefficients can be related to the collision geometry and density fluctuations of the colliding nuclei~\cite{Alver:2010gr,Alver:2010dn,ALICE:2011ab} and to the transport properties of the QGP in relativistic viscous hydrodynamic models~\cite{Gale:2012rq,Niemi:2015qia,Shen:2014vra,Bernhard:2016tnd,Bernhard2019}.

% TODO more pA measurements
The ridge structures in high-multiplicity pp and p--Pb events have been attributed to mechanisms like initial-state effects, such as gluon saturation~\cite{Dusling:2012cg,Bzdak:2013zma} and color reconnections~\cite{Ortiz:2013yxa,Sarma:2019teo} forming along the longitudinal direction and final-state effects, such as parton-induced interactions~\cite{Arbuzov:2011yr}, and collective effects due to a hydrodynamic behavior of the produced particles arising in a high-density system possibly formed in these collisions~\cite{Weller:2017tsr,Zhao:2017rgg}. 
Hybrid models implementing both effects are generally used in hydrodynamic simulations~\cite{Greif:2017bnr,Mantysaari:2017cni}. 
The proton shape and its fluctuations are also important to model the small systems \cite{Mantysaari:2017cni}.
The hydrodynamics itself might not be the only mechanism of the observed collectivity~\cite{Zhao:2020pty}. 
The influence and interplay of initial state and final state effects are recently studied carefully for the first time in ~\cite{Greif:2019ygb}, pointing out that the details of the initial state are crucially important for the quantitative description of observables in small systems~\cite{Schenke:2019pmk}. 
The attempts to describe the collective effects systematically from the small to large systems are being made experimentally~\cite{Acharya:2019vdf} and theoritically~\cite{Greif:2019ygb}.
However, a quantitative description of the full set of experimental data has not been achieved yet.
The summary of various explanations for the observed correlations in these small systems are summarized in ~\cite{Strickland:2018exs,Loizides:2016tew,Nagle:2018nvi}.

Furthermore, if collectivity in small systems is due to final state interactions, it is expected to measure its effect on jets. Proving the presence of jet quenching will be another crucial milestone to demonstrate the existence of the final-state effect in high-multiplicity pp collisions. The most of observables for the jet quenching in pp and p--Pb collisions do not show any clear evidences so far~\cite{Khachatryan:2016odn,Adam:2016jfp,Adam:2016dau,Acharya:2017okq}. The difficulties are attributed to an ambiguous reference since the hard probes themselves are also enhanced by requiring high multiplicity in the event~\cite{Adam:2016jfp,Acharya:2018egz}.

%Small system studies: A theory overview 1807.07191 Strickland:2018exs
%find one more one this regard.
The ATLAS experiment recently showed that the ridge remains in events tagged with a Z-boson~\cite{Aaboud:2019mcw}, possibly with an accompanying jet.
The impact parameter dependence on dijet or multi-jet production in pp collisions was studied in \cite{Frankfurt:2003td}.
The microscopic model for collectivity, based on interacting string implemented in $\pythiae$ event generator so called the ``String Shoving model''~\cite{Bierlich:2017vhg}, can qualitatively reproduce the CMS near-side ridge yield~\cite{Khachatryan:2016txc}. This challenges the hydrodynamic picture and predicts modifications to jet fragmentation
properties~\cite{Bierlich:2019ixq}.

@@

To further investigate these effects, we studied long-range correlations as a function of transverse momentum in very high multiplicity pp collisions at $\sqrt{s} =13$ TeV, collected with the high multiplicity event trigger during the LHC Run 2 data taking period. In this article, we present the near-side per-trigger yield at large pseudorapidity separation as a function of transverse momentum. The results are compared to previous measurements from CMS experiments. In addition, we report the ridge yield in events where harder fragmentation processes are present, to explore possible physical origins of long-range correlations.
The experimental setup and analysis method are described in Sec.~\ref{sec:experiment} and \ref{sec:ana}, sequentially.  The sources of systematic uncertainties are represented in Sec.~\ref{sec:uncertainties}. The results of the measurements and comparisons to model calculations are interpreted in Sec.~\ref{sec:results}. Finally, Sec.~\ref{sec:summary} summarizes our new results.

>>>>

Measurements of two-particle angular correlations are typically performed in terms of two dimensional $\Delta\eta$--$\Delta\varphi$ correlation functions, where $\eta$ is the pseudorapidity and $\varphi$ is the azimuthal angle. Long-range structure of two-particle correlation functions is of particular interest in studies of possible novel collective effects, where the effects of known sources such as resonance decays and fragmentation of high-momentum partons are known to be small. For most Monte Carlo (MC) event generators in pp collisions, the typical sources of such long-range correlations are momentum conservation and away-side ($\Delta\varphi$ $\approx$ $\pi$) jet correlations.
The enhancement in the associated yield of two-particle correlations at small relative azimuthal angle ($\Delta\varphi$) that extends over a long-range of relative pseudorapidity ($\Delta\eta$), is dubbed ``ridge'' due to its shape in the $\Delta\eta$--$\Delta\varphi$ plane.
The shape of these $\Delta\varphi$ correlations can be studied via a Fourier decomposition~\cite{Poskanzer:1998yz,Voloshin:2008dg}. The second and third order terms are the dominant harmonic coefficients $v_n$. The $v_n$ coefficients can be related to the collision geometry and density fluctuations of the colliding nuclei~\cite{Alver:2010gr,Alver:2010dn,ALICE:2011ab} and to the transport properties of the QGP in relativistic viscous hydrodynamic models~\cite{Gale:2012rq,Niemi:2015qia,Shen:2014vra,Bernhard:2016tnd,Bernhard2019}.

% TODO more pA measurements
The ridge structures in high-multiplicity pp and p--Pb events have been attributed to mechanisms from initial-state effects or final-state effects. The initial-state effects are usually referred to gluon saturation~\cite{Dusling:2012cg,Bzdak:2013zma} and color reconnections~\cite{Ortiz:2013yxa,Sarma:2019teo} forming along the longitudinal direction. The final-state effects are explained by such as parton-induced interactions~\cite{Arbuzov:2011yr} or collective effects due to a hydrodynamic behavior of the produced particles arising in a high-density system possibly formed in these collisions~\cite{Weller:2017tsr,Zhao:2017rgg}. 
Hybrid models implementing both effects are generally used in hydrodynamic simulations~\cite{Greif:2017bnr,Mantysaari:2017cni}. 
The proton shape and its fluctuations are also important to model the small systems~\cite{Mantysaari:2017cni}.
The hydrodynamics itself might not be the only mechanism of the observed collectivity~\cite{Zhao:2020pty}. 
The influence and interplay of the initial state and final state effects have recently been studied carefully for the first time in ~\cite{Greif:2019ygb}, pointing out that the details of the initial state are crucially important for the quantitative description of observables in small systems~\cite{Schenke:2019pmk}. 
The attempts to describe the collective effects systematically from small to large systems are being experimentally~\cite{Acharya:2019vdf} and theoritically~\cite{Greif:2019ygb} made.
However, a quantitative description of the full set of experimental data has not been achieved yet.
A summary of various explanations for the observed correlations in these small systems is given in ~\cite{Strickland:2018exs,Loizides:2016tew,Nagle:2018nvi}.

Furthermore, if collectivity in small systems is due to final state interactions, it is expected to have an effect on jets. Proving the presence of jet quenching will be another crucial milestone to demonstrate the existence of the final-state effect in high-multiplicity pp collisions. Most of the observables for the jet quenching in pp and p--Pb collisions do not show any clear evidence so far~\cite{Khachatryan:2016odn,Adam:2016jfp,Adam:2016dau,Acharya:2017okq}. The difficulties are attributed to an ambiguous reference since the hard probes themselves are also enhanced by requiring high multiplicity event~\cite{Adam:2016jfp,Acharya:2018egz}.
%Small system studies: A theory overview 1807.07191 Strickland:2018exs
%find one more one this regard.
The ATLAS experiment recently showed that the ridge remains in events tagged with a Z-boson~\cite{Aaboud:2019mcw}, possibly with an accompanying jet.
The impact parameter dependence on dijet or multi-jet production in pp collisions was studied in~\cite{Frankfurt:2003td,Frankfurt:2010ea}.
The microscopic model for collectivity, based on interacting string implemented in $\pythiae$ event generator as the so called the ``String Shoving model''~\cite{Bierlich:2017vhg}, can qualitatively reproduce the CMS near-side ridge yield~\cite{Khachatryan:2016txc}. This challenges the hydrodynamic picture and predicts modifications to jet fragmentation
properties~\cite{Bierlich:2019ixq}.

@@

To further investigate these effects, we studied long-range correlations as a function of transverse momentum in very high multiplicity pp collisions at $\sqrt{s} =13$ TeV, collected with the high multiplicity event trigger during the LHC Run 2 data taking period. In this article, we present the near-side per-trigger yield at large pseudorapidity separation as a function of transverse momentum. The results are compared to previous measurements by the CMS collaboration~\cite{Khachatryan:2015lva}. In addition, we report the ridge yield in events where harder fragmentation processes are present, to explore possible physical origins of the long-range correlations.
The experimental setup and analysis method are described in Sec.~\ref{sec:experiment} and \ref{sec:ana}, sequentially. The sources of systematic uncertainties are presented in Sec.~\ref{sec:uncertainties}. The results of the measurements and comparisons to model calculations are given in Sec.~\ref{sec:results}. Finally, Sec.~\ref{sec:summary} summarizes our new results.
